### 2018-08-26 paper/preliminaries.tex
First part of updates for complete version

In this section we present the background material for our work.

@@

% =========================================================================
\subsubsection*{CSP}
\label{section:CSP}

CSP is a process algebra for refinement. Using CSP, we model both systems and
their components using processes. There are characterised by their patterns
of interactions, modelled by synchronous, instantaneous, and atomic events.

@@

\begin{example}
  We consider the processes $P$, $Q$, and $R$ below. $P$ is initially
  ready to engage in the event $a$, and then makes a choice to
  behave as $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators allow us to define a choice over a
collection of processes $P(i)$.

@@

Interactions that are not supposed to be visible to the environment can be
hidden. The operator $P \ H$ defines a process that behaves as $P$, with the
interactions modelled by events in the set $H$ hidden. Frequently, hiding in
used in conjunction with parallelism, so that communications in a network of
parallel processes used just to coordinate them are affected
by the environment.

A rich collection of process operators allow us to define networks of
parallel processes in a concise and elegant way, and reason about safety,
liveness, and divergences.  A comprehensive account of the notation is given
in~\cite{Roscoe2010}.

A distinctive feature of CSP is its treatment of refinement~(as opposed to
bisimulation), which is convenient for reasoning about program correctness,
due to its treatment of nondeterminism and divergence.  A variety of semantic
models capture different notions of refinement. The simplest model
characterises a process by its possible traces:~sequence of events in which
it can engage.  We say that a process $P$ is trace-refined by another process
$Q$, written $P \lessdet_T Q$, if the set $traces(Q)$ of traces of $Q$ is
contained in the set $traces(P)$ of traces of $P$.

In fact, in every model, subset containment is used to define refinement. The
model we focus here first is the failures model, which captures both
sequences of interactions and deadlock behaviour. A failure of a process $P$
is a pair $(t,X)$ containing a trace $t$ of $P$ and a refusal:~a set $X$ of
events in which $P$ may refuse to engage, having engaged in the events of
$t$. The failures model of a process $P$ records all its failures in a set
$failures(P)$.

Failures refinement, $\lessdet_F$, is defined by set containment, but can
also be expressed in terms of traces refinement and another conformance
relation called $conf$ and defined below. The set $Ref(P,t)$ used in this
definition, and also specified below, contains the refusals of $P$ after the
trace $t$.
%
\begin{displaymath}
  Q\ conf\ P \defs \forall t: traces(P) \cap traces(Q) @ Ref(Q,t)
  \subseteq Ref(P,t)
  \\ %
  \mbox{where } Ref(P,t) \defs \{~X | (t,X) \in failures(P)~\}
\end{displaymath}
%
Conformance according to $conf$ requires that no new deadlocks are introduced
in an implementation. The theorem below, proved in~\cite{CG07}, establishes
that failures refinement amounts to traces refinement and $conf$.
%
\begin{theorem}
  $(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)$
  \label{thm:fref-tref-conf}
\end{theorem}
%
Process equivalence in each model is given simply by mutual refinement.

Throughout this paper, the alphabet of the processes, that is the set of
events that are in scope, is denoted by $\Sigma$. Since we are only
considering ``classical'' CSP processes, the alphabet is always finite. The
FDR tool~\cite{fdr} supports model checking and semantic analyses of
finite-state CSP processes.

@@

As shown in~\cite{Roscoe:1994:CME:197600}, any finite-state CSP process $P$
can be represented by a \emph{normalised transition graph}
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition from $n$ with label $e$. Normalisation of
$G(P)$ is reflected by the fact that $t$ is a function. The graph construction
in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable
by a sequence of edges labelled by $e_1\dots e_k$ and connecting states
$\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]
%
A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if and
only if there is a path through $G(P)$ starting  at $\ii n$ whose edge labels
coincide with $s$. The set of traces of $P$ is denoted by $\trc(P)$. We use
the analgous notation $\trc(G(P))$ for the set of finite, initialised paths
through $G(P)$, each path represented by its finite sequence of edge labels.
If $s\in\trc(P)$, then the process corresponding to $P$ after having executed
$s$ is denoted by $P/s$. Since $G(P)$ is normalised, there is a unique node
reached by applying the events from $s$ one by one, starting in $\ii n$.
Therefore, $G(P)/s$  is also well defined. By $[n]^0$ we denote the
\emph{initials} of $n$:~the set of events occurring as labels in any outgoing
transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
We also use this notation for CSP processes:~$[P]^0$ is the set of events $P$
may engage into, in other words, the initials of $P$ after the empty trace of
events, that is, $initials(P/\langle\rangle)$ as defined
in~\cite{Roscoe2010}, for example.

The total function $r$ maps each node $n$ to its \emph{refulsals} $r(n) =
\refs(n)$. Each element of $r(n)$ is a set of events that the CSP process $P$
might refuse to engage into, when in a process state corresponding to $n$.
 If $P/s$ is deterministic, its refusals coincide with the set of
subsets of $\Sigma - [P/s]^0$, including the empty set.

For finite CSP processes, since the refusals of each process state are
subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}, $\refs(P/s)$ can be
constructed from the set of \emph{maximal refusals}
$\maxrefs(P/s)\subseteq\refs(P/s)$. More formally, $\maxrefs(P/s)$ is defined
as follows.
%
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
%
Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
reconstruct the refusals $\refs(P/s)$ by subset-closure as follows.
%
\begin{equation}
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
%
The cardinality of $\maxrefs(P/s)$ reflects the degree of
nondeterminism that is present in process state $P/s$: the more maximal refusal sets
contained in  $\maxrefs(P/s)$, the more nondeterministic is the behaviour in
state $P/s$. Deterministic process states $P/s$ have exactly the one maximal refusal
$\Sigma-[P/s]^0$.

To see that this approach works only for finite CSP processes, we consider
the example where $\Sigma$ is infinite. In this case,
$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
defined above. As with refusals, we also use the transition graph-oriented
notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
associated with graph state $n$: if $n$ is the state reached in the
transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
= \maxrefs(P/s)$.

Well-formed normalised transition graphs must not refuse an event of the
initials of a state in {\it every} refusal applicable in this state; more
formally,
%
\begin{equation}
\label{eq:wellformedg}
\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
\exists R\in \maxrefs(n): e\not\in R
\end{equation}
%
By construction, normalised transition graphs reflect the \emph{failures
semantics} of finite-state CSP processes:~the traces $s$ of a process are
defined by the sequences of transitions associated with paths through its
graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
r(G(P)/s)$ represent the failures of $P$.

When investigating  tests for failures refinement, the notion of
\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
is also useful. An acceptance set of $P/s$ is a subset of the initials
$[P/s]^0$, i.e., a subset of events labelling  outgoing transitions of
$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
this set. If $P/s$ is nondeterministic, it internally chooses one of its
\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
refusing the events in $\Sigma-A$. The acceptances of $P/s$ are denoted by
$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
the following properties.
%
\begin{eqnarray}
A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
\label{eq:accref}
\\
\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
\label{eq:accinitials}
\\
 X\in\accs(P/s) &\Leftrightarrow & A\in \minaccs(P/s) \wedge A\subseteq X \subseteq [P/s]^0
 \label{eq:accsubset}
\end{eqnarray}
%
%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
%acceptances by means of maximal refusals. In this case,
%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
%sure? (4) does not talk about Acc. Is this important?}
%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
%removed it.}
%
Exploiting formulas (\ref{eq:accref}), (\ref{eq:accinitials}), and
(\ref{eq:accsubset}), every node of a normalised transition graph can
alternatively be labelled with their minimal acceptances, and this
information is equivalent to that contained in the maximal refusals. Since
process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
normalised transition graph, we also write $\minaccs(G(P)/s)$ and
$\maxrefs(G(P)/s)$ to denote the minimal acceptences and maximal refusals,
labelling graph state $G(P)/s$ and note that $\minaccs(G(P)/s) =
\minaccs(P/s)$ and $\maxrefs(G(P)/s) = \maxrefs(P/s)$.

@@

\begin{example}\label{ex:a}
Consider the CSP process $P$ in Example~\ref{example:CSP}. Its transition
graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state $P/\langle\rangle$
is represented there as Node\_0, with $\{ a\}$ as the only minimal
acceptance, since $a$ can never be refused, and no other events are accepted.
Having engaged into $a$, the transition emanating from Node\_0 leads to
Node\_2 representing  the process state $P/a = Q\intchoice R$. The internal
choice operator induces several minimal acceptances derived from $Q$ and $R$.
Since these processes accept their initial events in external choice,
$Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$ and $\{b,c\}$. Note
that event $c$ can never be refused, since it is contained in each minimal
acceptance set.

@@

Refinement relations between finite-state CSP processes $P, Q$ can be be
expressed by means of their normalised transition graphs in the following
way. %\fixme{alcc: trc was defined for P, not G(P) and this lemma is used
%later for trc(P). So, we need to be clearer as to what we mean.}
%\fxnote{jp: I have now introduced notation $\trc(G(P))$ above, so I think that the
%statement of the lemma is now sufficiently clear.}
%
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
  \\
  \label{eq:failrefa}
  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
  \\ & & \forall s\in\trc(G(Q)), R_Q\in\maxrefs(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in\maxrefs(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
  \\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s): \nonumber
  \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
  \end{eqnarray}
  \xbox
\end{lemma}
%
Traces refinement is, as already explained, characterised by subset inclusion
of sets of traces, here defined in terms of the normalised graph of a
process. In~(\ref{eq:failrefa}), we have a result similar to
Theorem~\ref{thm:fref-tref-conf}, but rely on the fact that refusal sets are
subset closed. In other words, if a process may refuse to engage in all
events in a set $X$, then it can refuse to engage in all events of subsets of
$X$. So, in considering $conf$, it is enough to consider maximal refusals.
Finally, (\ref{eq:failrefb}) captures the same result using acceptances,
rather than refusals, which as already observed, can be calculated by taking
the complement of refusals. So, acceptances of $Q$ are required to have a
corresponding larger set of acceptances in $P$.

@@

\subsubsection*{Tool Considerations}
FDR provides an API~\cite{fdrmanual} that can be used to construct normalised
transition graphs for CSP processes. The FDR graph nodes are labelled by
\emph{minimal acceptances} instead of maximal refusals as described above.
Since such a minimal acceptance set is the complement of a maximal refusal,
the function $r$ mapping states to their refusals can be implemented by
taking the complements of all minimal acceptances and then building all their
subsets. For practical applications, the subset closure is never constructed
in an explicit way; instead, sets are checked with respect to containment in
a maximal refusal.

@todo

@@

% =========================================================================
\subsection{Test Cases and Complete Test Suites}
\label{sec:cspcompletedef}
% =========================================================================

@todo

>>>>

@@

% =========================================================================
%\subsubsection*{CSP}
%\label{section:CSP}

\subsubsection*{Communicating Sequential Processes (CSP)} is a process algebra 
supporting system development by refinement. Using CSP, we model both systems and
their components using processes. There are characterised by their patterns
of interactions, modelled by synchronous, instantaneous, and atomic events.

@@

\begin{example}\label{example:CSP}
  We consider the processes $P$, $Q$, and $R$ below. $P$ is initially
  ready to engage in the event $a$, and then makes a choice to
  behave as $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators allow us to define a choice over a
collection of processes $P(i)$.

@@

Interactions that are not supposed to be visible to the environment can be
hidden. The operator $P \hide H$ defines a process that behaves as $P$, with the
interactions modelled by events in the set $H$ hidden. Frequently, hiding is used
in conjunction with parallelism: it may be desirable to make interal actions of each process in a network invisible, while events happening at their interfaces remain observable.

%Frequently, hiding is
%used in conjunction with parallelism, so that communications in a network of
%parallel processes used just to coordinate them are affected
%by the environment.

A rich collection of process operators allows us to define networks of
parallel processes in a concise and elegant way, and reason about safety,
liveness, and divergences.  A comprehensive account of the notation is given
in~\cite{Roscoe2010}.

A distinctive feature of CSP is its treatment of refinement~(as opposed to
bisimulation), which is convenient for reasoning about program correctness,
due to its treatment of nondeterminism and divergence.  A variety of semantic
models capture different notions of refinement. The simplest model
characterises a process by its possible \emph{traces}; the set $\trc(P)$ denotes
 the sequences of (non-hidden) 
events in which
$P$ can engage.  We say that a process \emph{$P$ is trace-refined by another process
$Q$}, written $P \lessdet_T Q$, if $\trc(Q)\subseteq \trc(P)$.

In fact, in every semantic model, subset containment is used to define refinement. The
model we focus here first is the failures model, which captures both
sequences of interactions and deadlock behaviour. A failure of a process $P$
is a pair $(s,X)$ containing a trace $s$ of $P$ and a refusal:~a set $X$ of
events in which $P$ may refuse to engage, after having performed the events of
$s$. The failures model of a process $P$ records all its failures in a set
$\fails(P)$. 
Using notation $P/s$ to denote the process $P$ after having engaged into trace $s$,
the set 
$\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains the \emph{refusals} 
of $P$ after the
trace $t$. 
Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure of $P$ and $Y\subseteq X$,
then $(s,Y)\in\fails(P)$ and $Y\in\refs(P/s)$ follows.

Failures refinement, $P \lessdet_F Q$, is defined by set containment 
$\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed, this implies
$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. As a consequence, failures 
refinement implies trace refinement. Therefore, using the additional conformance relation
\begin{equation}\label{eq:conf}
  Q\ conf\ P \defs \forall s\in \trc(P) \cap \trc(Q): \refs(Q/s)
  \subseteq \refs(P/s),
\end{equation}
failures refinement can be expressed by $\lessdet_T$ and $conf$.
\begin{equation}\label{thm:fref-tref-conf}
(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)
\end{equation}

%
%
%
%but can
%also be expressed in terms of traces refinement and another conformance
%relation called $conf$ and defined below. The set 
%$\refs(P,t) \defs \{~X | (t,X) \in \fails(P)~\}$ used in this
%definition contains the refusals of $P$ after the
%trace $t$.
%%
%\begin{displaymath}
%  Q\ conf\ P \defs \forall t: \trc(P) \cap \trc(Q) @ \refs(Q,t)
%  \subseteq \refs(P,t)
%\end{displaymath}
%%
%Conformance according to $conf$ requires that no new deadlocks are introduced
%in an implementation. The theorem below, proved in~\cite{CG07}, establishes
%that failures refinement amounts to traces refinement and $conf$.
%%
%\begin{theorem}
%  $(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)$
%  \label{thm:fref-tref-conf}
%\end{theorem}
%%
%Process equivalence in each model is given simply by mutual refinement.

Throughout this paper, the alphabet of the processes, that is the set of
events that are in scope, is denoted by $\Sigma$ and supposed to be finite. 
%Since we are only
%considering ``classical'' CSP processes, the alphabet is always finite. 
The
FDR tool~\cite{fdr} supports model checking and semantic analyses of
finite-state CSP processes.

For finite CSP processes, since the refusals of each process state are
subset-closed, $\refs(P/s)$ can be
constructed from the set of \emph{maximal refusals}; these are defined by 
%
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
%
Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
reconstruct the refusals $\refs(P/s)$ by  
%
\begin{equation}
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
%
The cardinality of $\maxrefs(P/s)$ reflects the degree of
nondeterminism that is present in process state $P/s$: the more maximal refusal sets
contained in  $\maxrefs(P/s)$, the more nondeterministic is the behaviour in
state $P/s$. Deterministic process states $P/s$ have exactly the one maximal refusal
$\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of $P/s$, i.e.,   the events that $P/s$ may engage into.

@@

As shown in~\cite{Roscoe:1994:CME:197600}, the failures semantics of 
any finite-state CSP process $P$
can be represented by a \emph{normalised transition graph}
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition, that is, an outgoing edge, from $n$ with label $e$, leading to node $t(n,e)$. Normalisation of
$G(P)$ is reflected by the fact that $t$ is a function. The graph construction
in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable
sequences of edges labelled by $e_1\dots e_k$ and connecting states
$\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]
%
By construction, $s\in\Sigma^*$ is a trace of $P$, if and only if 
there is a path through $G(P)$ starting  at $\ii n$ whose edge labels
coincide with $s$. In analogy to $\trc(P)$, we use
notation $\trc(G(P))$ for the set of finite, initialised paths
through $G(P)$, each path represented by its finite sequence of edge labels, and we
note that $\trc(P) = \trc(G(P))$. 
Since $G(P)$ is normalised, there is a unique node
reached by applying the events from $s$ one by one, starting in $\ii n$.
Therefore, $G(P)/s$  is also well defined. By $[n]^0$ we denote the
\emph{initials} of $n$:~the set of events occurring as labels in any outgoing
transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
The graph construction from~\cite{Roscoe:1994:CME:197600} guarantees that
$[G(P)/s]^0 = [P/s]^0$ for all traces of $P$.

The total function $r$ maps each node $n$ to a set of non-empty subsets of $\Sigma$. 
The graph construction guarantees that $r(G(P)/s)$ represents the maximal refusals
of $P/s$ for all 
$s\in\trc(P)$. As a consequence,  
\begin{equation}\label{eq:refsAndR}
(s,X)\in\fails(P) \iff s\in\trc(G(P)) \wedge \exists R\in r(G(P)/s): X\subseteq R,
\end{equation}
so $G(P)$ allows us to re-construct the failures semantics of $P$.

%
%
%To see that this approach works only for finite CSP processes, we consider
%the example where $\Sigma$ is infinite. In this case,
%$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
%calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
%defined above. As with refusals, we also use the transition graph-oriented
%notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
%associated with graph state $n$: if $n$ is the state reached in the
%transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
%= \maxrefs(P/s)$.

%Well-formed normalised transition graphs must not refuse an event of the
%initials of a state in {\it every} refusal applicable in this state; more
%formally,
%%
%\begin{equation}
%\label{eq:wellformedg}
%\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
%\exists R\in r(n): e\not\in R
%\end{equation}
%%
%
%
%By construction, normalised transition graphs reflect the \emph{failures
%semantics} of finite-state CSP processes:~the traces $s$ of a process are
%defined by the sequences of transitions associated with paths through its
%graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
%r(G(P)/s)$ represent the failures of $P$.

% =========================================================================
\subsubsection*{Acceptances}
\label{sec:accs}

When investigating  tests for failures refinement, the notion of
\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
is also useful. An acceptance set of $P/s$ is a subset of its initials
$[P/s]^0$; equivalently, it is a subset of events labelling  outgoing transitions of
$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
this set. If $P/s$ is nondeterministic, it internally chooses one of its
\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
refusing the events in $\Sigma-A$. The acceptances of $P/s$ are denoted by
$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
the following properties.
%
\begin{eqnarray}
A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
\label{eq:accref}
\\
\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
\label{eq:accinitials}
\\
 X\in\accs(P/s) &\Leftrightarrow & A\in \minaccs(P/s) \wedge A\subseteq X \subseteq [P/s]^0
 \label{eq:accsubset}
\end{eqnarray}
%
%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
%acceptances by means of maximal refusals. In this case,
%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
%sure? (4) does not talk about Acc. Is this important?}
%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
%removed it.}
%
Exploiting formulas (\ref{eq:accref}), (\ref{eq:accinitials}), and
(\ref{eq:accsubset}), every node of a normalised transition graph can
alternatively be labelled with their minimal acceptances, and this
information is equivalent to that contained in the maximal refusals. Since
process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
normalised transition graph, we also write $\minaccs(G(P)/s)$ and
note that (\ref{eq:refsAndR}) and (\ref{eq:accref}) imply 
\begin{equation}\label{eq:maxrefsminaccs}
\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \}.
\end{equation}

@@

\begin{example}\label{ex:a}
Consider the CSP process $P$ in Example~\ref{example:CSP}. Its transition
graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state $P/\varepsilon$ 
($\varepsilon$ denotes the empty trace)
is represented there as Node\_0, with $\{ a\}$ as the only minimal
acceptance, since $a$ can never be refused, and no other events are accepted.
Having engaged into $a$, the transition emanating from Node\_0 leads to
Node\_2 representing  the process state $P/a = Q\intchoice R$. The internal
choice operator induces several minimal acceptances derived from $Q$ and $R$.
Since these processes accept their initial events in external choice,
$Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$ and $\{b,c\}$. Note
that event $c$ can never be refused, since it is contained in each minimal
acceptance set.

@@

Summarising, refinement relations between finite-state CSP processes $P, Q$ can be be
expressed by means of their normalised transition graphs 
$$G(P) = ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))$$ 
and 
$$G(Q) = ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))$$
in the following
way. 
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
  \\
  \label{eq:failconf}
  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge P\ conf\ Q
  \\
  \label{eq:failrefa}
  P\ conf\ Q & \Leftrightarrow & \nonumber
  \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
   & \Leftrightarrow & 
    \forall s\in\trc(G(Q))\cap \trc(G(P)), A_Q\in\minaccs(G(Q)/s):  \nonumber
   \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
 \end{eqnarray}
  \xbox
\end{lemma}
%\begin{lemma}
%  \label{lemma:tgtrcref}
%  \begin{eqnarray}
%  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
%  \label{eq:trcrefa}
%  \\
%  \label{eq:failrefa}
%  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
%  \\ & & \forall s\in\trc(G(Q)), R_Q\in r_Q(G(Q)/s):  \nonumber
%  \\ & & \tabd
%  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
%  \\
%  \label{eq:failrefb}
%  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
%  \\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s): \nonumber
%  \\ & & \tabd
%  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
%  \end{eqnarray}
%  \xbox
%\end{lemma}
%

Formula (\ref{eq:trcrefa}) reflects trace refinement in terms of graph traces.  Formula~(\ref{eq:failrefa}) states how the $conf$ conformance relation can be expressed
by means of the maximal refusal functions of the graphs involved, and (\ref{eq:failrefb})
states the same in terms of the minimal acceptances that can be derived from the 
maximal refusal functions by means of (\ref{eq:maxrefsminaccs}).

%, we have a result similar to
%Theorem~\ref{thm:fref-tref-conf}, but rely on the fact that refusal sets are
%subset closed. In other words, if a process may refuse to engage in all
%events in a set $X$, then it can refuse to engage in all events of subsets of
%$X$. So, in considering $conf$, it is enough to consider maximal refusals.
%Finally, (\ref{eq:failrefb}) captures the same result using acceptances,
%rather than refusals, which as already observed, can be calculated by taking
%the complement of refusals. So, acceptances of $Q$ are required to have a
%corresponding smaller set of acceptances in $P$.

@@

%\subsubsection*{Tool Considerations}
%FDR provides an API~\cite{fdrmanual} that can be used to construct normalised
%transition graphs for CSP processes. The FDR graph nodes are labelled by
%\emph{minimal acceptances} instead of maximal refusals as described above.
%Since such a minimal acceptance set is the complement of a maximal refusal,
%the function $r$ mapping states to their refusals can be implemented by
%taking the complements of all minimal acceptances and then building all their
%subsets. For practical applications, the subset closure is never constructed
%in an explicit way; instead, sets are checked with respect to containment in
%a maximal refusal.
%
%@todo

@@

%% =========================================================================
%\subsection{Test Cases and Complete Test Suites}
%\label{sec:cspcompletedef}
%% =========================================================================
%
%@todo
